---
title: "Quick_Project_TA"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    
author: "Gloria & Vanessa"
date: "10/13/2019"
---

```{r}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

# Load Package and Data
```{r}
#load package 
library(gridExtra)
library(GGally)
library(tidyverse)
library(gridExtra)

#load data 
my_data <- read.csv("qp1_data.csv")
head(my_data)
fig_s<- "fig.width = 6, fig.height = 3"
```

```{r}
#add transit id
tran_id <- c(1:nrow(my_data))
my_data <- add_column(my_data, tran_id, .before = "price")
head(my_data)
```

# Base EDA
## Uni-variate non-graphical EDA

```{r}
str(my_data)
```

```{r}
summary(my_data)
```

```{r}
summary(my_data[,2:ncol(my_data)])
```


[Number of Bedrooms]
```{r}
#bedroom number count
table(my_data$bedrooms)
```

Note: Since bedroom number is generally the first buyers’ priority, we concluded that the highest demand goes to the house with three bedrooms type.

Question: with no bedroom? studio? a house with 33 bedrooms? may want to deal with it as an outlier.


```{r}
#pull out the 33 rooms entry
my_data%>%filter(my_data$bedrooms==33)
```

Note: It might be a typo since there are only 1620 sq ft for living and one bathroom if I have the authority I will remove this data.


[Number of floors]
```{r}
#floor numer count
table(my_data$floor)
```

[Transition in different zip code]
```{r}
table(my_data$zip_code)
```


## Uni-variate graphical EDA

[Price,sqft_live,sqft_lot Distribution]

```{r}
grid.arrange( 
  
ggplot(my_data, aes(x= price)) + geom_histogram(),
ggplot(my_data,aes(x=sqft_living))+geom_histogram(),
ggplot(my_data,aes(x=sqft_lot))+geom_histogram(),
ncol=1
)
```

Note: Price, sqft_living, sqft_lot distribution are all skewed. It's better use median instead of mean later. 

[floors, bedrooms, bathrooms Distribution]
```{r}
grid.arrange( 
  
ggplot(my_data, aes(x= floors)) + geom_histogram(stat="count"),
ggplot(my_data, aes(x=bedrooms)) + geom_histogram(stat="count"),
ggplot(my_data, aes(x=bathrooms)) + geom_histogram(stat="count"),
ncol=1
)
```

Finding:
Most of houses have 1 to 2 floors.
Most of houses have 3 to 4 rooms.
Most of houses have 1 or 2.5 bathrooms.


[House Location Distribution]

```{r,fig_s}
ggplot(my_data,aes(x=zip_code))+geom_bar()
```

Note: House distribution in each zone is not even.

[Bedroom Number Distribution]
```{r}
#just for checking the outlier house 
ggplot(my_data,aes(x=bedrooms))+geom_bar()+coord_cartesian(xlim=c(0,11))
```

Note: The market is looking for a house with three or four bedrooms.

[Year of Built Distribution]
```{r}
yr_built_pic <- ggplot(my_data,aes(x=yr_built))+geom_bar(stat = "count") +
theme( axis.text.x = element_text(size=10))
ggsave(filename = "yr_built.png",width = 8, height = 5, plot = yr_built_pic)
```

Note: House built from 2014 has highest number among King County area.

## multi-variate non-graphical EDA

[Price & zip_code]

```{r}
zip_price <- my_data %>%
  group_by(zip_code) %>%
  summarize(median(price, na.rm = TRUE))

colnames(zip_price)[2] <- "h_price"

zip_price
```

Note:House located in 98004/98039/98112	are pretty expensive, which can due to specific attribution with that area, a graph may demonstrate better.

[Bedroom & Bathroom]

```{r}
table(my_data$bedrooms,my_data$bathroom)
```

Note: more bedrooms with more bathrooms normally

*The 33 rooms house only has 1.75 bathroom


[Floor & View]
```{r}
table(my_data$floor,my_data$view)
```

Note: buildings with more floors are more likely with a better view


## Multi-variate graphical EDA

Which attribution affect the price most
```{r}
#Clean data for ploting 
plot_data <- data.frame(my_data)
plot_data$floors <- as.factor(plot_data$floors)
plot_data$bedrooms <- as.factor(plot_data$bedrooms)
plot_data$bathrooms <- as.factor(plot_data$bathrooms)
plot_data$view <- as.factor(plot_data$view)
plot_data$waterfront <- as.factor(plot_data$waterfront)
plot_data$grade <- as.factor(plot_data$grade)
plot_data$condition <- as.factor(plot_data$condition)

summary(plot_data)
```


```{r fig.width = 10, fig.height = 20}

grid.arrange(

ggplot(plot_data,aes(x=plot_data$bedrooms,y=price))+geom_boxplot(),
ggplot(plot_data,aes(x=plot_data$bathrooms,y=price))+geom_boxplot(),
ggplot(plot_data,aes(x=plot_data$sqft_living,y=price))+geom_point()+stat_smooth(method = "lm"),
ggplot(plot_data,aes(x=plot_data$sqft_lot,y=price))+geom_point()+stat_smooth(method = "lm"),
ggplot(plot_data,aes(x=plot_data$floors,y=price))+geom_boxplot(),
ggplot(plot_data,aes(x=plot_data$waterfront,y=price))+geom_boxplot(),
ggplot(plot_data,aes(x=plot_data$view,y=price))+geom_boxplot(),
ggplot(plot_data,aes(x=plot_data$grade,y=price))+geom_boxplot(),
ggplot(plot_data,aes(x=plot_data$condition,y=price))+geom_boxplot(),

ncol=2
)

```

Findings:
The relation between each factor and price are a bellow : 
-  A number of bedrooms increases, price increases, but the effect goes less significant when it's more than 5 bedrooms.
-  Number of bathroom increases, price increases,
-  Both spaces of living and lot have a positive relationship with price, but living space has more power than a lot space.
- Building with more floors at a higher price. However, the price reaches its peak at 2.5 floors and then drop.
-  The view doesn't really increase the value of the house. However, the waterfront view significantly did. 
-  The initial quality of the building is decisive to price, but the maintenance no so much.

[Price and Zip_code]
```{r, fig.width = 6, fig.height = 10}
#group the zip_code
zip_price <- my_data %>%
  group_by(zip_code) %>%
  summarize(median(price, na.rm = TRUE))

colnames(zip_price)[2] <- "h_price"
zip_price

```

```{r fig.width = 6, fig.height = 10}
#plotting 
ggplot(zip_price,aes( x=reorder(zip_code,h_price),y=h_price))+geom_bar(stat = "identity")+coord_flip()+ geom_hline(yintercept=median(my_data$price), color= "red")

```

Question : why the house located in 98039 98004 98040 98112 are more expensive than others?

[Prcie & space of living ]
```{r}
new_d <- my_data%>%filter(bedrooms<=8)
new_d$bedrooms <- as.factor(new_d$bedrooms)
ggplot(new_d,aes(x=sqft_living , y= price,color=bedrooms))+geom_point(alpha = 0.5,size=1)+stat_smooth(method="lm",se=FALSE)
```

Findings :
-  More sqft_living increase the price, this phenomenon will be more significant if there are more rooms in the house
- But the effect of sqft to price drops when there 7 rooms in the house, may due to the over-divided interior design
- The effect bounces back when house has 8 rooms, but there are only 18 observations of this group. I don't really trust this 

```{r}
#pull out all house with 8 room to check if there are other factor that may affect the price 
summary(my_data %>% filter(bedrooms==8))
```

Finding :
The house with 8 rooms is more expenxive even the living space doesn't increase may because this group is comparatively higher in grade and condition. However, interestingly, they barely come with a view or waterfront view, which intuitionally will cause the price to decrease. 

[Price and year built]

```{r}
library(tidyverse)
plot_data %>% group_by(yr_built) %>% summarize(med_pri = median(price)) %>% ggplot(mapping = aes(x = yr_built, y = med_pri)) + geom_line()

```

Finding: The relation comes in a U shape, maybe people like the old house for its vintage style and the new one for its condition.

[Price Trend & Floors]
!! Premise !!! The transaction record is chronological 

```{r}
ggplot(my_data[1:5000,],aes(x=tran_id, y=price))+geom_line()+facet_grid(floors~.)

```

Finding: The price change pattern is not the same in different bedrooms group.


# Findings (Executive Advice )
## Main sales force city

Tartget: Find out which city and what kind of house occupied the most sale

```{r fig.width = 6, fig.height = 10}
# group by zip_code, but take sum of the price for sales force analysis
zip_price_sum <- my_data %>%
  group_by(zip_code) %>%
  summarize(sum(price, na.rm = TRUE))

colnames(zip_price_sum)[2] <- "h_price"
zip_price_sum <- zip_price_sum%>%arrange(-h_price)

# Let's look at the zip_code price sum 
ggplot(zip_price_sum,aes( x=reorder(zip_code,h_price),y=h_price))+geom_bar(stat = "identity")+coord_flip()
```

```{R}
# try to find which city sales the most. Intuitionally, Seattle should occupy the most sales.
# 98102-98199 - Seattle 
Seattle_sales<- my_data %>% filter(zip_code%in% c(98102,98199))%>%summarise(sum(price))
Seattle_market_share <- as.double(Seattle_sales / sum(my_data$price))


#We will also check some city from the top
#98004-98009 bellevue
bellevue_sales<- my_data %>% filter(zip_code%in%c(98004,98009))%>%summarise(sum(price))
bellevue_market_share <- as.double(bellevue_sales / sum(my_data$price))

#98040 Mercer island
mercer_sales<- my_data %>% filter(zip_code==98040)%>%summarise(sum(price))
mercer_market_share <- as.double(mercer_sales/sum(my_data$price))

#98052 53 Redmond
Redmond_sales<- my_data %>% filter(zip_code%in%c(98052,98053))%>%summarise(sum(price))
Redmond_market_share <- as.double(Redmond_sales/sum(my_data$price))

# 98033 98034 kirkland
kirkland_sales<- my_data %>% filter(zip_code%in%c(98033,98034))%>%summarise(sum(price))
kirkland_market_share <- as.double(kirkland_sales/sum(my_data$price))
kirkland_market_share
```


```{r, fig.width = 5, fig.height =3}
#gather data for plotting
mkt_share <- c(Seattle_market_share,bellevue_market_share,mercer_market_share,Redmond_market_share,kirkland_market_share)
mkt_city <- c("Seattle","Bellevue","Mercer Island","Redmond","kirkland")
mkt <- data.frame(mkt_city,mkt_share )

#plotting
mkt_share <- ggplot(mkt, aes(x= mkt_city,y=mkt_share))+geom_bar(stat = "identity")+labs(title = "Main Sales Force Market Share", x="City",y ="Market Share")+ geom_text(aes(label=round(mkt_share,2)),color = "DarkGreen",vjust=-0.10)+theme_classic()
mkt_share

ggsave("mkt_share.png")
```

Finding: Redmond makes most of the money, followed by Kirkland.

```{r}
#Now let check what type of houese these buyers are looking for
# We will foucusing on the more relevant factor based on previous analysis: bedrooms, sqrt_living and  
# Redmon
Redmon_sales<- plot_data%>%filter(zip_code%in%c(98052,98053))
summary(Redmon_sales)

```
```{r}
#plotting for choice preference 
ggplot(Redmon_sales, aes(y=sqft_living ,x=grade, color= bedrooms))+geom_point(alpha=0.2)+ geom_hline(yintercept=quantile(Redmon_sales$sqft_living),probs=.25,color= "red",linetype="dashed", size=0.5)+ geom_jitter()+labs(title = "Redmon Customer Preference", x="Grade",y ="Living Space(Sqrt)",color="Bedrooms")

ggsave("Remond_mkt.png")
```

Finding:
Poeple in Redmon are mainly looking for house with 3 or 4 bedrooms. around 2350 sqrt, and grade in 8.

```{r}
#Kirkland
Kirkland_sales<- plot_data%>%filter(zip_code%in%c(98033,98034))
summary(Kirkland_sales)
```

```{r}
#plotting for choice preference 
ggplot(Kirkland_sales, aes(y=sqft_living ,x=grade, color= bedrooms))+geom_point(alpha=0.2)+ geom_hline(yintercept=quantile(Kirkland_sales$sqft_living),probs=.25,color= "red",linetype="dashed", size=0.5)+ geom_jitter()+labs(title = "Kirkland Customer Preference", x="Grade",y ="Living Space(Sqrt)",color="Bedrooms")

ggsave("Kirkland_mkt.png")
```

Finding:
Poeple in Kirkland are mainly looking for house with 3 or 4 bedrooms. around 2000 sqrt, and grade in 7


## The power of rennovation

Tartget: Find out how much value the renovation can bring to the price 
```{r}
# We use price per sqft as idex, for ruling out the effect of house size
t_data <- data.frame(my_data)
t_data$sqrt_price <- t_data$price/ t_data$sqft_living
#head(t_data)

# add one more column
for(i in 21613){
  t_data$renovated <- NA
}


# put in data

t_data$renovation <- ifelse(t_data$yr_renovated ==0, "N", "Y")
head(t_data)
```


```{r fig.width = 15, fig.height = 50}
#plot the price increse 
t_data$yr_built <- as.factor(t_data$yr_built)

ggplot(t_data,aes(x=yr_built,y=price,color = renovation))+geom_boxplot()+coord_flip()
```

Note: There's a real difference in price the renovation can bring. But we try to focus more on how much the difference 


```{r}
# prepare data 
reno_power <- t_data%>%group_by(renovation,yr_built)%>%summarise(median(sqrt_price))
colnames(reno_power)[3] <- "price"
head(reno_power)
```

```{r fig.width = 16 , fig.height = 10}
#plot the difference
reno_pic <- ggplot(reno_power,aes(x = yr_built, y=price, group = renovation)) + 
geom_line(aes(color=renovation), size = 2.0) +
labs(fill = "Renovation", size=10.0) +
ggtitle("The difference between renovation and without renovation") +
ylab("Median of Price per sqft") +
labs(fill = "Renovation") +
theme(plot.title = element_text(face="bold", size=20)) +
theme(axis.title = element_text(face="bold", size=17), axis.text.x = element_text(size=6,angle=45), axis.text.y = element_text(size=10))
```

Finding: The price of of renovated houese( blue line) is almost always higher the those without(red). And sometime is times of difference.

```{r}
ggsave(filename = "reno_pic.png", width = 16, height = 10, plot = reno_pic)
```




